(working title) Local Physiological Adaptation in \textit{Mimulus cardinalis}

### 2015-04-22 ms/ms.Rnw
Worked on manuscript Results

@@

  # Plot latitude versus PC1
  pdf(paste(pathFig, "/Figure_PC1vLat.pdf", sep = ""), 5, 5)
  par(mar = c(5, 5, 1, 1))
  plot(focClim$Lat, focClim$PC1, xlim = c(32, 44), ylim = c(-4, 4), axes = F,
    frame.plot = T, type = "n", xlab = "", ylab = "")
  title(xlab = "Latitude", cex.lab = 1.5)
  title(ylab = substitute(plain(Principal~Component~1)~group("(", x, ")"), 
    list(x = pve)))
  axis(1, at = seq(32, 44, 4), labels = c(expression(32*degree), expression(36*degree),
    expression(40*degree), expression(44*degree)), lwd = 0, lwd.ticks = 1) 
  axis(2, at = seq(-4, 4, 2), las = 1, lwd = 0, lwd.ticks = 1)    
  fitLat <- lm(PC1 ~ Latitude, data = focClim)

@@

@@

@@

  # Columns with climatic variables to test
  load(paste(pathDat, "/Exploratory_climate_variables.RData", sep = "")) #climVars

  # Variable selection using random forests (VSURF)
  linmod.PC1 <- as.formula(paste("PC1 ~ ", paste(climVars, collapse = " + "), sep = ""))
  set.seed(69646)
  vsurfPC1 <- VSURF.parallel(linmod.PC1, data = focClim[, c("PC1", climVars)], 
    ncores = 3)

@@

\begin{table}[ht]
   \centering
   \topcaption{Climatic variables}
   \begin{tabular}{@{} ll @{}}
      \toprule
  Climate variable & Abbreviation \\
      \midrule
  % climateWNA VARIABLES:
	DD\_0	& degree-days below 0\celsius (chilling degree-days)  \\
	DD5	  & degree-days above 5\celsius (growing degree-days)   \\
	DD\_18	& degree-days below 18\celsius (heating degree-days)  \\
	DD18	& degree-days above 18\celsius (cooling degree-days)  \\
	NFFD	& number of frost-free days                           \\
	PAS	  & precipitation as snow (mm) between August in previous year and July in current  \\
	Eref	& Hargreaves reference evaporation (mm) \\
	CMD	  & Hargreaves climativ moisture deficit (mm) \\
	MAR   &	mean annual solar radiation (NOTE: removing because too many missing values) \\
	RH    &	mean annual relative humidity \\
  %BIOCLIM VARIABLES:
  bio1	&	Annual Mean Temperature	\\
	bio2	&	Mean Diurnal Range (Mean of monthly (max temp - min temp))	\\
	bio3	&	Isothermality (bio2/bio7) (* 100)	\\
	bio4	&	Temperature Seasonality (standard deviation *100)	\\
	bio5	&	Max Temperature of Warmest Month	\\
	bio6	&	Min Temperature of Coldest Month	\\
	bio7	&	Temperature Annual Range (bio5-bio6)	\\
	bio8	&	Mean Temperature of Wettest Quarter	\\
	bio9	&	Mean Temperature of Driest Quarter	\\
	bio10	&	Mean Temperature of Warmest Quarter	\\
	bio11	&	Mean Temperature of Coldest Quarter	\\
	bio12	&	Annual Precipitation	\\
	bio15	&	Precipitation Seasonality (Coefficient of Variation)	\\
	bio16	&	Precipitation of Wettest Quarter	\\
	bio17	&	Precipitation of Driest Quarter	\\
	bio18	&	Precipitation of Warmest Quarter	\\
	bio19	&	Precipitation of Coldest Quarter	\\
	    \bottomrule
   \end{tabular}
\end{table}

@@

\begin{table}[ht]
   \centering
   \topcaption{Population, treatment effects}
   \begin{tabular}{@{} llllllll @{}}
      \toprule
                                   & Germination & Leaf expansion & Stem elongation & Photosynthesis & Intrin. Photo & Mortality \\
                                   & survreg     & lmer           & lmer            & lmer           & lmer          & glm \\
      \midrule
  Population                       &
    \Sexpr{sigStar(anova(fitGerm)["Pr(>Chi)"]["PopID",])}  &
    \Sexpr{sigStar(fitLER$anova.table["PopID", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID", "Pr(>F)"])} & 
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID", "Pr(>Chisq)"])} & \\
  Temperature                      &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["TempTrt", "Pr(>F)"])}    &
    \Sexpr{sigStar(fitSER$anova.table["TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["TempTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["TempTrt", "Pr(>F)"])} &     \Sexpr{sigStar(Anova(fitMort, type = 2)["TempTrt", "Pr(>Chisq)"])} & \\
  Water                            &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["WaterTrt", "Pr(>F)"])}   &
    \Sexpr{sigStar(fitSER$anova.table["WaterTrt", "Pr(>F)"])}   &
    \Sexpr{sigStar(fitPhoto$anova.table["WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["WaterTrt", "Pr(>Chisq)"])} & \\
  Pop $\times$ Temp                &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["PopID:TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID:TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID:TempTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID:TempTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID:TempTrt", "Pr(>Chisq)"])} & \\
  Pop $\times$ Water               &
     NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["PopID:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID:WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID:WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID:WaterTrt", "Pr(>Chisq)"])} & \\
  Temp $\times$ Water              &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["TempTrt:WaterTrt", "Pr(>Chisq)"])} & \\
  Pop $\times$ Temp $\times$ Water &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID:TempTrt:WaterTrt", "Pr(>Chisq)"])} & \\
      \bottomrule
   \end{tabular}
\end{table}

@@

\subsection*{Little evidence population by treatment interactions}

@@

% ANOVA table for LLL growth rate (need to label tables)
\begin{table}[htbp]
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[ANOVA table, leaf expansion rate]{CAPTION}
	\begin{center}
	\begin{tabular}{lcccccc}
	\toprule

% ANOVA table for height growth rate (need to label tables)
\begin{table}[htbp]
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[ANOVA table, stem expansion rate]{CAPTION}
	\begin{center}
	\begin{tabular}{lcccccc}
	\toprule

\subsection*{Latitude versus growth rate}

@@

@@

>>>>

\begin{figure}[h!]
	\centerline{\includegraphics[width=1\textwidth]{Figures/Figure_ExptlDes.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Experimental Design]{CAPTION}
	\label{fig:Fig_ExptlDes}
\end{figure}

@@

  # Plot latitude versus PC1
  pdf(paste(pathFig, "/Figure_PC1vLat.pdf", sep = ""), 5.5, 5)
  par(mai = c(1, 1.5, 0.25, 0.25))
  plot(focClim$Lat, focClim$PC1, xlim = c(32, 44), ylim = c(-4, 4), axes = F,
    frame.plot = T, type = "n", xlab = "", ylab = "")
  title(xlab = "Latitude", cex.lab = 1.5)
  title(ylab = substitute(plain(Principal~Component~1)~group("(", x, ")"), 
    list(x = pve)))
  axis(1, at = seq(32, 44, 4), labels = c(expression(32*degree), 
		expression(36*degree), expression(40*degree), expression(44*degree)), 
		lwd = 0, lwd.ticks = 1) 
  axis(2, at = seq(-4, 4, 2), las = 1, lwd = 0, lwd.ticks = 1)    
  fitLat <- lm(PC1 ~ Latitude, data = focClim)

	# Label y-axis from slow to fast growth
	mtext(c("Slow\nGrowth", "Fast\nGrowth"), 2, at = c(-4, 4), cex = 1.25, 
				font = 3, las = 1, line = 4, adj = 0.5)

@@

As with previous studies in \textit{Mimulus} and many other species [CITES], latitude \textit{per se} predicted trait variation more than many obvious environmental variables (e.g. temperature or precipitation) that covary with latitude.

@@

	# Load climVars
	load(paste(pathDat, "/ExploratoryCimateVariables.RData", sep = ""))

@@

  # Variable selection using random forests (VSURF)
  linmod.PC1 <- as.formula(paste("PC1 ~ ", paste(climVars, collapse = " + "), sep = ""))
  set.seed(69647)
  vsurfPC1 <- VSURF.parallel(linmod.PC1, data = focClim[, c("PC1", climVars)], 
    ncores = 3)

@@

\begin{table}[ht]
   \centering
   \topcaption{Climatic variables}
   \begin{tabular}{@{} ll @{}}
      \toprule
  Climate variable & Abbreviation \\
      \midrule
  % climateWNA VARIABLES:
	DD\_0	& degree-days below 0\celsius (chilling degree-days)  \\
	DD5	  & degree-days above 5\celsius (growing degree-days)   \\
	DD\_18	& degree-days below 18\celsius (heating degree-days)  \\
	DD18	& degree-days above 18\celsius (cooling degree-days)  \\
	NFFD	& number of frost-free days                           \\
	PAS	  & precipitation as snow (mm) between August in previous year and July in current  \\
	Eref	& Hargreaves reference evaporation (mm) \\
	CMD	  & Hargreaves climativ moisture deficit (mm) \\
	MAR   &	mean annual solar radiation (NOTE: removing because too many missing values) \\
	RH    &	mean annual relative humidity \\
  % BIOCLIM VARIABLES:
  bio1	&	Annual Mean Temperature	\\
	bio2	&	Mean Diurnal Range (Mean of monthly (max temp - min temp))	\\
	bio3	&	Isothermality (bio2/bio7) (* 100)	\\
	bio4	&	Temperature Seasonality (standard deviation *100)	\\
	bio5	&	Max Temperature of Warmest Month	\\
	bio6	&	Min Temperature of Coldest Month	\\
	bio7	&	Temperature Annual Range (bio5-bio6)	\\
	bio8	&	Mean Temperature of Wettest Quarter	\\
	bio9	&	Mean Temperature of Driest Quarter	\\
	bio10	&	Mean Temperature of Warmest Quarter	\\
	bio11	&	Mean Temperature of Coldest Quarter	\\
	bio12	&	Annual Precipitation	\\
	bio15	&	Precipitation Seasonality (Coefficient of Variation)	\\
	bio16	&	Precipitation of Wettest Quarter	\\
	bio17	&	Precipitation of Driest Quarter	\\
	bio18	&	Precipitation of Warmest Quarter	\\
	bio19	&	Precipitation of Coldest Quarter	\\
	    \bottomrule
   \end{tabular}
\end{table}

%--------------------------------------------------
% Results
%--------------------------------------------------

@@

@@

Genetic differences between populations affecting originsmal function and performance point to traits that may be involved in local adaptation. Using a common garden design, we identified strong genetic differences in germination, growth, and photosynthetic rate among populations of \textit{M. cardinalis} (Table~\ref{table:Table_AnovaSummary}). A single principal component captured \Sexpr{pve}$\%$ of the trait variation among populations, defining an axis of variation from fast to slow growth (Fig~\ref{fig:Fig_PC1vLat}). As we explain below, intrinsic differences between populations in terms of plant function (photosynthesis) and performance (growth) contrasted greatly with little variation in plasticity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of Population, Treatment, Population x Treatment interactions for varies traits
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[ht]
   \centering
   \topcaption{Population, treatment effects}
   \begin{tabular}{@{} llllllll @{}}
      \toprule
                                   & Germination & Leaf expansion & Stem elongation & Photosynthesis & Intrin. Photo & Mortality \\
                                   & survreg     & lmer           & lmer            & lmer           & lmer          & glm \\
      \midrule
  Population                       &
    \Sexpr{sigStar(anova(fitGerm)["Pr(>Chi)"]["PopID",])}  &
    \Sexpr{sigStar(fitLER$anova.table["PopID", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID", "Pr(>F)"])} & 
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID", "Pr(>Chisq)"])} & \\
  Temperature                      &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["TempTrt", "Pr(>F)"])}    &
    \Sexpr{sigStar(fitSER$anova.table["TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["TempTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(Anova(fitMort, type = 2)["TempTrt", "Pr(>Chisq)"])} & \\
  Water                            &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["WaterTrt", "Pr(>F)"])}   &
    \Sexpr{sigStar(fitSER$anova.table["WaterTrt", "Pr(>F)"])}   &
    \Sexpr{sigStar(fitPhoto$anova.table["WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["WaterTrt", "Pr(>Chisq)"])} & \\
  Pop $\times$ Temp                &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["PopID:TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID:TempTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID:TempTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID:TempTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID:TempTrt", "Pr(>Chisq)"])} & \\
  Pop $\times$ Water               &
     NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["PopID:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID:WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID:WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID:WaterTrt", "Pr(>Chisq)"])} & \\
  Temp $\times$ Water              &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["TempTrt:WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["TempTrt:WaterTrt", "Pr(>Chisq)"])} & \\
  Pop $\times$ Temp $\times$ Water &
    NA                                                          &
    \Sexpr{sigStar(fitLER$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitSER$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} &
    \Sexpr{sigStar(fitPhoto$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} & 
    \Sexpr{sigStar(fitIntrnPhoto$anova.table["PopID:TempTrt:WaterTrt", "Pr(>F)"])} &     
    \Sexpr{sigStar(Anova(fitMort, type = 2)["PopID:TempTrt:WaterTrt", "Pr(>Chisq)"])} & \\
      \bottomrule
   \end{tabular}
   \label{table:Table_AnovaSummary}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure of latitude versus PC1 (growth axis)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
	\centerline{\includegraphics[width=1\textwidth]{Figures/Figure_PC1vLat.pdf}}
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[Southern populations grow faster]{CAPTION}
	\label{fig:Fig_PC1vLat}
\end{figure}

\subsection*{Remarkably little evidence for variation in plasticity}

Genotype $\times$ environment interactions are also a common signature of local adaptation [CITE]. For example, populations from more stressful environments may tradeoff reduced growth rate under benign conditions for the ability to tolerate stress by maintaining positive growth rates and/or surviving through adverse conditions. We found remarkably little evidence for this pattern in \textit{M. cardinalis}. There was only one statistically significant Population $\times$ Treatment interaction (Table~\ref{table:Table_AnovaSummary}), but this idiosyncratic result was weak and would not have survived correction for multiple testing. Otherwise, populations responded remarkably similarly to treatments: faster growth in the hot treatment, slower growth in the dry treatment, and high mortality in the hot, dry treatment (Table~\ref{table:Table_AnovaSummary}). Note that interactions were calculated after factoring out intrinsic trait differences, necessarily reducing statistical power to detect significant interactions relative to main effects. However, the fact that the Population and Treatment effects were highly significant ($P \lle 0.001$ in most cases) suggests that statistical power alone cannot explain why we failed to detect Population $\times$ Treatment interactions.

@@

Latitude explains variation
Presumably selective agent is not latitude, but something related to it
Exploratory approach to identify putative selective agent to test in future studies
We used a machine learning approach...identified interannual variation in precipitation
Obviously, cannot identify alternative variables that were not included in analysis.

%--------------------------------------------------
% Discussion
%--------------------------------------------------

@@

1. How does local adaptation occur? It's complex (gene flow, shape of tradeoffs, demography) and although there is abundant evidence for local adaptaiton, we do not a general sense for when local adaptation succeeds or fails. For this, we need a system in which there are axes of variation and constraint, traits with known physiological function?, and selective agents

2. Why does local adaptation succeed? shallow environmental gradients lead to convex tradeoffs not heavily constrained by gene flow.

3. Selective agent - unpredictability. Explain

4. Why does local adaptation fail? concave, high gene flow, demography. Qualify that our metrics of plasticity are coarse, but at least suggest that GxE is much weaker than G for the environmental gradients evaluated here.

5. Future work and broad conclusions. Supports general conclusions that traits related to timing and growth are evolutionarily labile whereas those related to the fundamental niche are constrained (cites like \cite{Emery_etal_2012, Emery_Ackerly_2014}).

@@

% ANOVA table for LER (need to label tables)
\begin{table}[htbp]
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[ANOVA table, leaf expansion rate]{CAPTION}
	\begin{center}
	\begin{tabular}{lcccccc}
	\toprule

% ANOVA table for SER (need to label tables)
\begin{table}[htbp]
	\usefont{T1}{phv}{m}{n}
	\fontsize{10}{12}
	\selectfont
	\caption[ANOVA table, stem expansion rate]{CAPTION}
	\begin{center}
	\begin{tabular}{lcccccc}
	\toprule

% Figures showing latitudinal clines with specific traits
